Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-23 article.tex
wording

Formally, the connectome can be represented as a directed graph $G=(V,E)$,
where $V$ is a set of $p$ nodes, representing neurons, and $E \subseteq
\left\{(i, j) \in V \times V\right\}$ is a set of edges, representing direct
synaptic connections between neurons. Causal interactions are expressed by the
direction of edges: $(i, j) \in E$ indicates that the state of neuron $j$ might
be caused by the activity of neuron $i$. In those terms,  the connectome
inference problem is formally stated as follows:  \textit{Given the sampled
observations $\{ x^t_i \in \mathbb{R}^{p} | i \in V, t = 1, \dots T \}$ of $p$
neurons for $T$ time intervals, the goal is to infer the set $E$ of connections in $G$.}

In this paper, we present a simplified -- yet nearly as good -- version of the
winning method\footnote{Code available at \url{https://github.com/asutera
/kaggle-connectomics}} of the Connectomics
Challenge\footnote{\url{http://connectomics.chalearn.org}}, as a simple and
theoretically grounded approach based on signal processing techniques and
partial correlation statistics. The paper is structured as follows: Section
\ref{sec:filter} describes the signal processing methods applied on fluorescent
calcium time series; Section \ref{sec:inference} then presents the proposed
approach and its theoretical properties; Section \ref{sec:results} provides an
empirical analysis and comparison with other network inference methods, while
Section \ref{sec:conclusion} finally discusses our work and provides further
research directions. Additionally, Appendix \ref{app:optimized} further
describes, in all details, our actual winning method, giving slightly better
results than the method presented in the paper but highly tuned for the
challenge.

>>>>

Formally, the connectome can be represented as a directed graph $G=(V,E)$,
where $V$ is a set of $p$ nodes, representing neurons, and $E \subseteq
\left\{(i, j) \in V \times V\right\}$ is a set of edges, representing direct
synaptic connections between neurons. Causal interactions are expressed by the
direction of edges: $(i, j) \in E$ indicates that the state of neuron $j$ might
be caused by the activity of neuron $i$. In those terms,  the connectome
inference problem is formally stated as follows:  \textit{Given the sampled
observations $\{ x^t_i \in \mathbb{R} | i \in V, t = 1, \dots, T \}$ of $p$
neurons for $T$ time intervals, the goal is to infer the set $E$ of connections in $G$.}

In this paper, we present a simplified -- yet nearly as good -- version of the
winning method\footnote{Code available at \url{https://github.com/asutera
/kaggle-connectomics}} of the Connectomics
Challenge\footnote{\url{http://connectomics.chalearn.org}}, as a simple and
theoretically grounded approach based on signal processing techniques and
partial correlation statistics. The paper is structured as follows: Section
\ref{sec:filter} describes the signal processing methods applied on fluorescent
calcium time series; Section \ref{sec:inference} then presents the proposed
approach and its theoretical properties; Section \ref{sec:results} provides an
empirical analysis and comparison with other network inference methods, while
Section \ref{sec:conclusion} finally discusses our work and provides further
research directions. Additionally, Appendix \ref{app:optimized} further
describes, in all details, our actual winning method, giving slightly better
results than the method presented in the paper at a price of significant
tuning.
